\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-16 2-the-cubble-pkg.Rmd
flesh out more in the beginning of sec 2 and compatibility

\textcolor{red}{more beginning} In cubble data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes. The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site. Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.

@@

To use functions in the `cubble` package, an analyst will need to first turn the data object into a \code{cubble} class. \code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude. The use of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form.

Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains 5 Australia weather stations, a subset of hundreds of weather stations in Australia. This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020. \code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:

@@

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch(ts)
cubble_long
```

@@

```{r}
attr(cubble_long, "spatial")
```

@@

The last 4 subsections introduces operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf}, and \code{netcdf4}.

@@

Building from an \code{tbl_df} class, cubble has implemented methods for \pkg{dplyr} generics, which includes:

  - basic: mutate, filter, summarise, select, arrange, rename, left_join
  - grouping: group_by, ungroup
  - slice family: slice_head, slice_tail, slice_sample, slice_min, slice_max

@@

Cubble is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, only the \code{coords} argument needs to be specified:

```{r echo = TRUE}
# casting a tsibble into cubble
dt <- climate_flat %>% 
  tsibble::as_tsibble(key = id, index = date) %>% 
  cubble::as_cubble(coords = c(long, lat))
dt
```

In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and existing operations on the tsibble class can still be applied:

@@

Cubble is compatible with simple feature geometry list-column (\code{sfc}) in \pkg{sf} and  \code{s2_lnglat} vector in \pkg{S2} in the nested form. After creating a cubble, analysts can add the geometry vectors as needed to perform spatial operations:

```{r echo = TRUE}
library(s2)
# create a cubble
cb <- climate_flat %>% 
  cubble::as_cubble(key = id, index = date, coords = c(long, lat))

# western australia map
wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")

# create a `s2_lnglat` vector based on long and lat and filter stations in wa
cb %>%
  mutate(ll = s2_lnglat(long, lat)) %>% 
  filter(s2_within(ll, wa))
```

```{r echo = TRUE}
library(sf)
# simplify australia map
aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)

dt <- cb %>%
  mutate(
    # create `sfc` column based on long and lat
    ll = st_sfc(
      purrr::map2(long, lat, ~st_point(c(.x, .y))),
      crs = st_crs(aus)),
    # append state multi-polygon based on the `sfc` created
    state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
    # adopt a different projection: lambert conformal conic (EPSG:3112)
    state = st_transform(state, crs = "EPSG:3112"))

@@

>>>>

In this section, the two formats cubble uses to arrange spatio-temporal data will be presented Then operations will be introduced, with short examples, to illustrate how to work with these two formats. Lastly, a subsection will be dedicated to how existing packages, in spatial and temporal analysis, fit in with cubble.

In cubble, data can represented in two formats: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms with the associated attributes. The decision on which form to use in data manipulation is output-oriented, meaning analysts need to first think about whether the output of an operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of time invariant variables, or an operation that summarises time varying variables into site. Underneath the nested form, a cubble is built from a \code{rowwise_df} class where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{purrr::map()} syntax when working with list-column.

@@

To use functions in the `cubble` package, an analyst will need to first turn the data object into a \code{cubble} class. \code{as_cubble()} does this through supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude and latitude. The use of \code{key} and \code{index} follows the design in the \pkg{tsibble} package. The cubble created by default is in the nested form.

Before the formal examples in Section \ref{examples}, each function introduced in this section will be demonstrated with a short example with a data that contains five Australia weather stations, a subset of hundreds of weather stations in Australia. This dataset contains spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information: date, maximum and minimum temperature and precipitation for 2020. \code{climate_flat} stores this data in format 1) in Figure \ref{fig:illu-input} and the code below creates a cubble out of \code{climate_flat} with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:

@@

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch()
cubble_long
```

@@

```{r echo = TRUE}
attr(cubble_long, "spatial")
```

@@

The last four subsections introduce operations specific to the \code{cubble} class and this section will demonstrate how the \code{cubble} class interacts with existing packages, specifically, \code{dplyr}, \code{tsibble}, \code{sf} )\code{s2}), and \code{netcdf4}.

@@

The \code{dplyr} package has provided many tools for common data wrangling tasks and these operations are useful in the spatio-temporal context. \code{cubble} provides its method for the following \code{dplyr} verbs to wrangle data in both the nested and long form:

 > `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_head`, `slice_tail`, `slice_sample`, `slice_min`, `slice_max`)

\textcolor{red}{more work on groupby} The grouping pair, \code{group_by} and \code{ungroup}, can only be used in the long form to define grouping on the time but not in the nested form.

@@

Tsibble is a temporal data structure that uses \code{index} and \code{key} to identify the time and different series. Cubble can be seen as a natural extension of tsibble for spatio-temporal data with an additional coordinates component. The use of two forms avoids the duplicates of spatial variables at each time stamp and hence, in the visualisation, avoid drawing duplicated points with the same coordinate. To cast a \code{tsibble} into a \code{cubble}, only the \code{coords} argument needs to be supplied:

```{r echo = TRUE}
# demonstrate with a tsibble created from climate_flat
raw <- climate_flat %>% 
  tsibble::as_tsibble(key = id, index = date)

dt <-  raw %>% 
  cubble::as_cubble(coords = c(long, lat))
dt
```

In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class and operations available to the tsibble class is still valid under cubble. For example, the code below calculate two features of the maximum temperature:

@@

As a spatial data object, \code{sf} introduces a simple feature geometry list-column (\code{sfc}) to the data frame and provides spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON} etc). While these operations are valueable for spatio-temporal data analysis, an \code{sf} object usually don't contain temporal variables and hence can't be directly cast into a \code{cubble}. However, \code{cubble} can contain \code{sfc} columns in the nested form and spatial operations built on the \code{sfc} class are available. The following example shows how to create an \code{sfc} column of \code{POINT} type based on latitude and longitude in cubble. The state \code{MULTIPOLYGON} is then added based on spatial operation \code{st_within} before making a coordinate transformation.

```{r echo = TRUE}
library(sf)
# create a cubble
cb <- climate_flat %>% 
  cubble::as_cubble(key = id, index = date, coords = c(long, lat))

aus <- ozmaps::abs_ste

dt <- cb %>%
  mutate(
    # create `sfc` column based on long and lat
    ll = st_sfc(
      purrr::map2(long, lat, ~st_point(c(.x, .y))),
      crs = st_crs(aus)),

    # append state multi-polygon based on the `sfc` created
    state = aus$geometry[st_within(ll, aus, sparse = FALSE)],

    # adopt a different projection: lambert conformal conic (EPSG:3112)
    state = st_transform(state, crs = "EPSG:3112")
    )

@@

An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for using \code{s2}-prefixed functions:

```{r echo = TRUE}
library(s2)
# western australia map
wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")

# mutate a `s2_lnglat` vector on `cb` created in the last chunk
cb %>%
  mutate(ll = s2_lnglat(long, lat)) %>% 
  filter(s2_within(ll, wa))
```

### 2022-02-16 main.tex
flesh out more in the beginning of sec 2 and compatibility

To use functions in the \texttt{cubble} package, an analyst will need to
first turn the data object into a \code{cubble} class.
\code{as_cubble()} does this through supplying the three key components:
\code{key} as the spatial identifier; \code{index} as the temporal
identifier; and a vector of \code{coords} in the order of longitude and
latitude. The use of \code{key} and \code{index} follows the convention
in the \pkg{tsibble} package. The cubble created by default is in the
nested form.

Before the formal examples in Section \ref{examples}, each function
introduced in this section will be demonstrated with a short example
with a data that contains 5 Australia weather stations, a subset of
hundreds of weather stations in Australia. This dataset contains spatial
information of each station: station id, latitude, longitude, elevation,
station name and World Meteorology Organisation ID and also daily
temporal information: date, maximum and minimum temperature and
precipitation for 2020. \code{climate_flat} stores this data in format
1) in Figure \ref{fig:illu-input} and the code below creates a cubble
out of \code{climate_flat} with \code{id} as the key, \code{date} as the
index, and \code{c(long, lat)} as the coordinates:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% stretch(ts)
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 5 x 6
  id            lat  long  elev name           wmo_id
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610
2 ASN00010311 -31.9  117. 179   york            94623
3 ASN00010614 -32.9  117. 338   narrogin        94627
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
5 ASN00015131 -17.6  134. 220   elliott         94236
\end{CodeOutput}
\end{CodeChunk}

@@

Building from an \code{tbl_df} class, cubble has implemented methods for
\pkg{dplyr} generics, which includes:

\begin{itemize}
\tightlist
\item
  basic: mutate, filter, summarise, select, arrange, rename, left\_join
\item
  grouping: group\_by, ungroup
\item
  slice family: slice\_head, slice\_tail, slice\_sample, slice\_min,
  slice\_max
\end{itemize}

@@

Cubble is also compatible with \pkg{tsibble}. When creating a cubble
from a tsibble, only the \code{coords} argument needs to be specified:

\begin{CodeChunk}
\begin{CodeInput}
R> # casting a tsibble into cubble
R> dt <- climate_flat %>% 
+   tsibble::as_tsibble(key = id, index = date) %>% 
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

In the created nested form, each element of the list-column \code{ts} is
of \code{tbl_ts} class and existing operations on the tsibble class can
still be applied:

@@

Cubble is compatible with simple feature geometry list-column
(\code{sfc}) in \pkg{sf} and \code{s2_lnglat} vector in \pkg{S2} in the
nested form. After creating a cubble, analysts can add the geometry
vectors as needed to perform spatial operations:

\begin{CodeChunk}
\begin{CodeInput}
R> library(s2)
R> # create a cubble
R> cb <- climate_flat %>% 
+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
R> 
R> # western australia map
R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
R> 
R> # create a `s2_lnglat` vector based on long and lat and filter stations in wa
R> cb %>%
+   mutate(ll = s2_lnglat(long, lat)) %>% 
+   filter(s2_within(ll, wa))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [3]: nested form
# bbox:     [115.97, -32.94, 117.18, -31.89]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts                 ll      
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> library(sf)
R> # simplify australia map
R> aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
R> 
R> dt <- cb %>%
+   mutate(
+     # create `sfc` column based on long and lat
+     ll = st_sfc(
+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
+       crs = st_crs(aus)),
+     # append state multi-polygon based on the `sfc` created
+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+     # adopt a different projection: lambert conformal conic (EPSG:3112)
+     state = st_transform(state, crs = "EPSG:3112"))
R> 
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name         wmo_id ts                        ll
  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <list>           <POINT [°]>
1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 <tibble~ (115.9764 -31.9275)
2 ASN00010311 -31.9  117. 179   york          94623 <tibble~  (116.765 -31.8997)
3 ASN00010614 -32.9  117. 338   narrogin      94627 <tibble~ (117.1797 -32.9342)
4 ASN00014015 -12.4  131.  30.4 darwin airp~  94120 <tibble~ (130.8925 -12.4239)
5 ASN00015131 -17.6  134. 220   elliott       94236 <tibble~ (133.5407 -17.5521)
# ... with 1 more variable: state <MULTIPOLYGON [m]>
\end{CodeOutput}
\end{CodeChunk}

>>>>

To use functions in the \texttt{cubble} package, an analyst will need to
first turn the data object into a \code{cubble} class.
\code{as_cubble()} does this through supplying the three key components:
\code{key} as the spatial identifier; \code{index} as the temporal
identifier; and a vector of \code{coords} in the order of longitude and
latitude. The use of \code{key} and \code{index} follows the design in
the \pkg{tsibble} package. The cubble created by default is in the
nested form.

Before the formal examples in Section \ref{examples}, each function
introduced in this section will be demonstrated with a short example
with a data that contains five Australia weather stations, a subset of
hundreds of weather stations in Australia. This dataset contains spatial
information of each station: station id, latitude, longitude, elevation,
station name and World Meteorology Organisation ID and also daily
temporal information: date, maximum and minimum temperature and
precipitation for 2020. \code{climate_flat} stores this data in format
1) in Figure \ref{fig:illu-input} and the code below creates a cubble
out of \code{climate_flat} with \code{id} as the key, \code{date} as the
index, and \code{c(long, lat)} as the coordinates:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% stretch()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> attr(cubble_long, "spatial")
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 5 x 6
  id            lat  long  elev name           wmo_id
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610
2 ASN00010311 -31.9  117. 179   york            94623
3 ASN00010614 -32.9  117. 338   narrogin        94627
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120
5 ASN00015131 -17.6  134. 220   elliott         94236
\end{CodeOutput}
\end{CodeChunk}

@@

The \code{dplyr} package has provided many tools for common data
wrangling tasks and these operations are useful in the spatio-temporal
context. \code{cubble} provides its method for the following
\code{dplyr} verbs to wrangle data in both the nested and long form:

\begin{quote}
\texttt{mutate}, \texttt{filter}, \texttt{summarise}, \texttt{select},
\texttt{arrange}, \texttt{rename}, \texttt{left\_join}, and the slice
family (\texttt{slice\_head}, \texttt{slice\_tail},
\texttt{slice\_sample}, \texttt{slice\_min}, \texttt{slice\_max})
\end{quote}

\textcolor{red}{more work on groupby} The grouping pair, \code{group_by}
and \code{ungroup}, can only be used in the long form to define grouping
on the time but not in the nested form.

@@

Tsibble is a temporal data structure that uses \code{index} and
\code{key} to identify the time and different series. Cubble can be seen
as a natural extension of tsibble for spatio-temporal data with an
additional coordinates component. The use of two forms avoids the
duplicates of spatial variables at each time stamp and hence, in the
visualisation, avoid drawing duplicated points with the same coordinate.
To cast a \code{tsibble} into a \code{cubble}, only the \code{coords}
argument needs to be supplied:

\begin{CodeChunk}
\begin{CodeInput}
R> # demonstrate with a tsibble created from climate_flat
R> raw <- climate_flat %>% 
+   tsibble::as_tsibble(key = id, index = date) 
R> 
R> dt <-  raw %>% 
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

In the created nested form, each element of the list-column \code{ts} is
of \code{tbl_ts} class and operations available to the tsibble class is
still valid under cubble. For example, the code below calculate two
features of the maximum temperature:

@@

As a spatial data object, \code{sf} introduces a simple feature geometry
list-column (\code{sfc}) to the data frame and provides spatial
operations on various geometry types (\code{POINT}, \code{LINESTRING},
\code{POLYGON}, \code{MULTIPOLYGON} etc). While these operations are
valueable for spatio-temporal data analysis, an \code{sf} object usually
don't contain temporal variables and hence can't be directly cast into a
\code{cubble}. However, \code{cubble} can contain \code{sfc} columns in
the nested form and spatial operations built on the \code{sfc} class are
available. The following example shows how to create an \code{sfc}
column of \code{POINT} type based on latitude and longitude in cubble.
The state \code{MULTIPOLYGON} is then added based on spatial operation
\code{st_within} before making a coordinate transformation.

\begin{CodeChunk}
\begin{CodeInput}
R> library(sf)
R> # create a cubble
R> cb <- climate_flat %>% 
+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
R> 
R> aus <- ozmaps::abs_ste
R> 
R> dt <- cb %>%
+   mutate(
+     # create `sfc` column based on long and lat
+     ll = st_sfc(
+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
+       crs = st_crs(aus)),
+     
+     # append state multi-polygon based on the `sfc` created
+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+     
+     # adopt a different projection: lambert conformal conic (EPSG:3112)
+     state = st_transform(state, crs = "EPSG:3112")
+     )
R> 
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name         wmo_id ts                        ll
  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <list>           <POINT [°]>
1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 <tibble~ (115.9764 -31.9275)
2 ASN00010311 -31.9  117. 179   york          94623 <tibble~  (116.765 -31.8997)
3 ASN00010614 -32.9  117. 338   narrogin      94627 <tibble~ (117.1797 -32.9342)
4 ASN00014015 -12.4  131.  30.4 darwin airp~  94120 <tibble~ (130.8925 -12.4239)
5 ASN00015131 -17.6  134. 220   elliott       94236 <tibble~ (133.5407 -17.5521)
# ... with 1 more variable: state <MULTIPOLYGON [m]>
\end{CodeOutput}
\end{CodeChunk}

An \pkg{s2} \code{lnglat} vector can similarly be created in cubble for
using \code{s2}-prefixed functions:

\begin{CodeChunk}
\begin{CodeInput}
R> library(s2)
R> # western australia map
R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
R> 
R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
R> cb %>%
+   mutate(ll = s2_lnglat(long, lat)) %>% 
+   filter(s2_within(ll, wa))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [3]: nested form
# bbox:     [115.97, -32.94, 117.18, -31.89]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts                 ll      
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
\end{CodeOutput}
\end{CodeChunk}
